---
title: "SQUEAC Attack"
author: "Mark Myatt"
date: 32 July 2018 #'`r format(Sys.Date(), "%d %B %Y")`'
fontsize: 12pt
geometry: margin=2cm
documentclass: article
classoption: a4paper
bibliography: bibliography.bib
link-citations: TRUE
links-as-notes: FALSE
colorlinks: TRUE
linkcolor: blue
citecolor: blue
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

work <- read.csv("squeacAttack.csv")

if(!require(knitr)) install.packages("knitr")
if(!require(kableExtra)) install.packages("kableExtra")
if(!require(magrittr)) install.packages("magrittr")

```

\newpage

# Calculate 'k' for single coverage estimate {#k}

Correction factor `k` is the ratio of the mean length of an untreated episode to the mean length of a CMAM treatment episode.

&nbsp;

$$ k ~ = ~ \frac{\text{Mean length of an untreated episode}}{\text{Mean length of a successful treatment episode}} $$

&nbsp;

Mean length of an untreated episode for SAM or MAM can be assumed as 7.5 months based on @Garenne:2009fq. Mean length of a successful treatment episode can be estimated from routine programming monitoring data by calculating the median length of stay.

This can be implemented in R as follows:

&nbsp;

```{r echo = TRUE, eval = TRUE}
medianLOS <- median(work$medianLOS, na.rm = TRUE)

k <- ((7.5 * 30.44) / 7) / work$medianLOS
```

&nbsp;

The **median of the median length of stay** of the surveys in the dataset is `r medianLOS`.

The `k` values for each of the surveys in the dataset are[^1]:

&nbsp;

```{r echo = FALSE, eval = TRUE}
k
```

\newpage

# Calculate r.out (recovering cases NOT in the program) {#r.out}

Using the calculated `k` values in previous section, number of recovering cases NOT in the program is calculated using the following formula:

&nbsp;

$$\begin{aligned}
R_{out} & ~ \approx ~ \frac{1}{k} ~ \times ~ \left ( ~ R_{in} ~ \times ~ \frac{C_{in} ~ + ~ C_{out} ~ + ~ 1}{C_{in} ~ + ~ 1} ~ - ~ R_{in} ~ \right ) \\
\\
where: & \\
\\
k & ~ = ~ \text{correction factor} \\
C_{in} & ~ = ~ \text{current SAM cases in the program} \\
C_{out} & ~ = ~ \text{current SAM cases not in the program} \\ 
R_{in} & ~ = ~ \text{recovering SAM cases in the program}
\end{aligned}$$

&nbsp;

This can be implemented in R as follows:

&nbsp;

```{r, echo = TRUE, eval = TRUE}
r.out <- floor((1 / k) * 
                 (work$r.in * ((work$c.in + work$c.out + 1) / 
                                 (work$c.in + 1)) - work$r.in))
```

&nbsp;

The resulting vector of `r.out` values are:

&nbsp;

```{r, echo = FALSE, eval = TRUE}
r.out
```

\newpage

# Calculate Prior modes (priorMode) from priorAlpha and priorBeta with their standard errors (priorModeSE)

The Prior mode can be calculated from Prior $\alpha$ and Prior $\beta$ as shown in the following formula.

&nbsp;

$$ mode_{Prior} ~ = ~ \frac{\alpha_{Prior} - 1}{\alpha_{Prior} ~ + ~ \beta_{Prior} ~ - ~ 2} $$

&nbsp;

This can be implemented in R as follows:

&nbsp;

```{r echo = TRUE, eval = TRUE}
priorN <- work$priorAlpha - 1

priorD <- work$priorAlpha + work$priorBeta - 2

priorMode <- priorN / priorD
```

&nbsp;

which results in:

&nbsp;

```{r priortable, echo = FALSE, eval = TRUE}
priortable <- data.frame(priorN, priorD, round(priorMode, digits = 2))

kable(x = priortable[1:20, ],
      caption = "Prior numerator, Prior denominator and Prior mode (first 20 records)",
      booktabs = TRUE,
      row.names = FALSE,
      col.names = c("$Prior_{numerator}$", "$Prior_{denominator}$", "$Prior_{mode}$"),
      escape = FALSE,
      format = "latex") %>%
  row_spec(row = 0, bold = TRUE) %>%
  kable_styling(latex_options = c("HOLD_position", "striped"),
                position = "center")
```

\newpage

# Calculate appropriate likelihood numerators (likeN) denominators (likeD)

The likelihood mode can be calculated depending on the coverage estimator to assess: *point coverage* or *period coverage*.

&nbsp;

$$\begin{aligned}
\text{Point coverage} ~ & = ~ \frac{C_{in}}{C_{in} + C_{out}} \\
\\
\text{Period coverage} ~ & = ~ \frac{C_{in} ~ + ~ R_{in}}{C_{in} ~ + ~ C_{out} ~ + ~ R_{in} ~ + ~ R_{out}}\\
\\
where: & \\
\\
C_{in} & ~ = ~ \text{current SAM cases in the program} \\
C_{out} & ~ = ~ \text{current SAM cases not in the program} \\ 
R_{in} & ~ = ~ \text{recovering SAM cases in the program} \\
R_{out} & ~ = ~ \text{recovering SAM cases not in the program}
\end{aligned}$$

&nbsp;

This can be implemented in R as follows:

&nbsp;

```{r, echo = TRUE, eval = TRUE}
likeN <- ifelse(work$coverType == "point", work$c.in, work$c.in + work$r.in)

likeD <- ifelse(work$coverType == "point", work$c.in + work$c.out, 
           ifelse(work$coverType == "period", 
                  work$c.in + work$r.in + work$c.out, 
                  work$c.in + work$r.in + work$c.out + r.out))

likeMode <- likeN / likeD
```

&nbsp;

which results in:

```{r liketable, echo = FALSE, eval = TRUE}
liketable <- data.frame(likeN, likeD, round(likeMode, digits = 2))

kable(x = liketable[1:20, ],
      caption = "Likelihood numerator, likelihood denominator and likelihood mode",
      booktabs = TRUE,
      row.names = FALSE,
      col.names = c("$Likelihood_{numerator}$", "$Likelihood_{denominator}$", "$Likelihood_{mode}$"),
      escape = FALSE,
      format = "latex") %>%
  row_spec(row = 0, bold = TRUE) %>%
  kable_styling(latex_options = c("HOLD_position", "striped"),
                position = "center")
```

\newpage

# Make summary data.frame

```{r, echo = TRUE, eval = TRUE}
results <- data.frame(priorN, priorD, priorMode, likeN, likeD, likeMode)

results <- results[!is.na(results$priorMode) & !is.na(results$likeMode), ]
```

The resulting data.frame is:

```{r resultsDF, echo = FALSE, eval = TRUE}
kable(x = results[1:30, ],
      caption = "Summary data.frame (first 30 records)",
      booktabs = TRUE,
      row.names = FALSE,
      col.names = c("$Prior_{numerator}$", "$Prior_{denominator}$", "$Prior_{mode}$", 
                    "$Likelihood_{numerator}$", "$Likelihood_{denominator}$", "$Likelihood_{mode}$"),
      escape = FALSE,
      format = "latex") %>%
  row_spec(row = 0, bold = TRUE) %>%
  kable_styling(latex_options = c("HOLD_position", "striped", "scale_down"),
                position = "center")
```

\newpage

# Test for  prior-likelihood conflict

```{r, echo = TRUE, eval = TRUE}
for(i in 1:nrow(results)) {
	# Make a two-by-two table
	tab <- matrix(c(results$priorN[i], results$priorD[i] - results$priorN[i],
	                results$likeN[i], results$likeD[i] - results$likeN[i]), 
	              nrow = 2, byrow = TRUE)
  # Fisher test (works with expected numbers < 5)
  results$p[i] <- round(fisher.test(round(tab))$p.value, 4)
}

results$PLC <- ifelse(results$p < 0.05, TRUE, FALSE)
```

```{r plctable, echo = FALSE, eval = TRUE}
kable(x = results[1:30, ],
      caption = "Summary data.frame with prior-likelihood conflict variable (first 30 records)",
      booktabs = TRUE,
      row.names = FALSE,
      col.names = c("$Prior_{numerator}$", "$Prior_{denominator}$", "$Prior_{mode}$", 
                    "$Likelihood_{numerator}$", "$Likelihood_{denominator}$", "$Likelihood_{mode}$",
                    "p-value", "PLC"),
      escape = FALSE,
      format = "latex") %>%
  row_spec(row = 0, bold = TRUE) %>%
  kable_styling(latex_options = c("HOLD_position", "striped", "scale_down"),
                position = "center")
```

\newpage

# How common are prior-likelihood conflicts?

```{r, echo = TRUE, eval = TRUE}
table(results$PLC)
round(prop.table(table(results$PLC)) * 100, 2)
```

\newpage

# How are the prior and likelihood modes related?

&nbsp;

```{r likehist, echo = TRUE, eval = TRUE, fig.cap = "Prior-Likelihood Relationship", fig.align = "center", fig.pos = "H", fig.width = 6, fig.height = 6, fig.retina = 1}
## Errors (difference)
error <- results$priorMode * 100 - results$likeMode * 100
summary(error)
hist(error, 
     breaks = 16, 
     xlab = "Prior mode (%) - Likelihood mode (%)", 
     ylab = "Number of assessments", 
     main = "")
```

\newpage

# Scatterplot

```{r likescatter, echo = TRUE, eval = TRUE, fig.cap = "Prior-Likelihood Relationship", fig.align = "center", fig.pos = "H", fig.width = 6, fig.height = 6, fig.retina = 1}
plot(results$priorMode * 100, 
     results$likeMode * 100, 
     xlim = c(0, 100), ylim = c(0, 100), 
     xlab = "Prior mode (%)", 
     ylab = "Likelihood mode (%)", 
     pch = ifelse(results$PLC, 19, 1), 
     frame.plot = FALSE)
abline(a = 0, b = 1, lty = 2)
text(100, 15, "Prior mode > Likelihood mode", pos = 2, cex = 0.8)
text(  0, 85, "Prior mode < Likelihood mode", pos = 4, cex = 0.8)
lines(lowess(results$priorMode * 100, results$likeMode * 100, f = 2/3))
cor(results$priorMode, results$likeMode)
```

\newpage

# How precise is the likelihood estimate alone

For this, we assume total population of 100,000 with 17% aged 6-59 months and prevalence of SAM of 2%.

```{r, echo = TRUE, eval = TRUE}
# Subset to results with PLC == TRUE
rejected <- results[results$PLC, ]
# calculate number of SAM
pop <- 100000 * 0.17 * 0.02 
# calculate finite population correction factor
rejected$FPC <- sqrt((pop - rejected$likeD) / (pop - 1))
```

## Relative precision

```{r, echo = TRUE, eval = TRUE}
# Relative precision of surveys with PLC
rejected$likeRP <- (qnorm(0.975) * 
                      sqrt((rejected$likeMode * 
                              (1 - rejected$likeMode)) / rejected$likeD) * 
                    rejected$FPC) / rejected$likeMode

# Relative precision of an EPI coverage survey with p = likelihood mode, 
# n = 120, and DEFF = 2.0?
rejected$epiRP <- (2.0 * qnorm(0.975) * 
                     sqrt((rejected$likeMode * 
                             (1 - rejected$likeMode)) / 210)) / rejected$likeMode

# How many have relative precision of better than or equal to the 
# assumed EPI survey?

table(rejected$likeRP <= rejected$epiRP)
prop.table(table(rejected$likeRP <= rejected$epiRP))
```

\newpage

```{r, echo = TRUE, eval = TRUE}
# Proportion of SQUEAC assessments that fail by ...
#
#    prior likelihood conflict == TRUE AND precision worse than the 
#    assumed EPI survey
#
failN <- sum(ifelse(rejected$likeRP > rejected$epiRP, 1, 0))
failP <- round(sum(ifelse(rejected$likeRP > rejected$epiRP, 1, 0)) / 
                 nrow(results) * 100, 2) 

print(failN)
print(failP)
```

\newpage

# References {#references}


[^1]: `NA` values are for surveys that don't report a median length of stay.